\begin{versionhistory}
  \vhEntry{0.5}{06.07.18}{{Gilles~Jacobs}}{Create draft with incomplete typology.}
  \vhEntry{0.6}{12.07.18}{{Gilles~Jacobs,~Diane Breesch,~Véronique Hoste,~Els Lefever}}{Discuss event typology with finance domain-expert: Remove \type{Stock Holding} from typology because it is too rare an event.}
  \vhEntry{0.7}{02.08.18}{{Gilles~Jacobs}}{Finish typology. Define event subtypes and participant definitions.}
  \vhEntry{0.8}{13.08.18}{{Gilles~Jacobs,~Véronique Hoste}}{Revise after review. Allow discontiguous spans which were previously disallowed. Refine typology. FILLER arguments are fully described in the typology. This version is used for annotator training and the first annotation test-run.}
  \vhEntry{0.9}{23.08.18}{{Gilles~Jacobs,~Diane Breesch}}{Add rule clarifications for annotator difficulties based on the annotation test-run. Majorly revise typology based on prof. Breesch domain-expert comments: Remove \type{CapitalReturns} type and \type{CapitalReturns\_SharePurchase} subtype: Share repurchases are exceptionally rare in the corpus so it is pruned. This removes the need for the \type{CapitalReturns} main type. Add a \type{Dividend} type with corresponding subtypes previously as \type{CapitalReturns\_DividendX}. Add \type{Expense} type and subtypes: expenses are often mentioned in corpus and together with \type{Revenue} and \type{Profit/Loss} form the three major factors in accounting. Renamed \type{FinancialResult} to \type{FinancialReport}: Specify that this event exclusively covers financial reports or trends in financial results of a company. Remove \type{Employment\_Problem} because it is too broad and rare.}
  \vhEntry{1.0}{03.09.18}{Gilles Jacobs}{Clarify conceptual overlap and how to differentiate between event types. Clarify rules and add some examples. Add \type{In/DecreaseAmount} participant on \type{Revenue} subtypes.} This is the version used for the annotation and correction of the v1.0 corpus.
  \vhEntry{1.1}{10.01.19}{Gilles Jacobs}{Add Guidelines Enhancement Proposal chapter. Fix typos and added more examples from corpus.}

\end{versionhistory}